AFQ-Insight

### 2019-12-06 reports/manuscript/sections/methods.tex
Convert to docx using pandoc. Simplify some of the math used to allow that.

\item Diffusion MRI data from a previous study of properties of the white matter
across the lifespan\cite{yeatman2014lifespan}, containing dMRI data from 76
subjects with ages 6-50. These data were measured in a GE Discovery 750 3T MRI
scanner at the Stanford Center for Cognitive and Neurobiological Imaging. The
Stanford University IRB approved the procedures of this study. Informed consent
was obtained from each adult participant, and assent for participation was
provided by parents/guardians for children. Voxel resolution was
\num{2x2x2}\si{\milli\meter^3} with 96 non-colinear directions measured with a
$b= \SI{2000}{\second\per\milli\meter^2}$ and 30 non-colinear directions
measured with a $b=\SI{1000}{\second\per\milli\meter^2}$.
Data was preprocessed to correct for subject motion and the diffusion
tensor model \cite{basser1994mr} was fit in every voxel, using a robust fit
\cite{chang2005restore}. These data were acquired using a dual spin echo
sequence, in which there is sufficient time for eddy currents to subside between
the application of the gradients and the image acquisition, so no eddy current
correction was appliced. We will refer to this dataset as WH.

\item Diffusion MRI from from a previous study of the corticospinal tract
profile in patients with amyotrophic lateral sclerosis
(ALS)\cite{sarica2017corticospinal}, containing data from 24 ALS patients and 24
demographically matched healthy controls. These data were measured in a GE
Discovery 750 3T MRI scanner at the Institute of Bioimaging and Molecular
Physiology in Catanzaro. Informed consent was provided as approved by the
Ethical Committee of the University ``Magna Graecia'' of Catanzaro. Voxel
resolution was \num{2x2x2}\si{\milli\meter^3} and 27 non-colinear directions
were measured with a $b=\SI{1000}{\second\per\milli\meter^2}$.
Here, both motion correction and eddy current correction were
applied before modeling the diffusioon tensor in each voxel. We will refer to
this dataset as ALS.

@@

Here, we propose that in addition to regularizing the coefficients in $\beta$,
we can also capitalize on our knowledge of the group structure of the tract
profile features in $\mathbf{X}$. The tract-metric combinations form a natural
grouping. For example, we expect that MD features within the left arcuate
fasciculus will co-vary across individuals. Likewise for FA values within the
right corticospinal tract (CST) and so on. This group structure is represented
in Fig~\ref{fig:group-structure}, which depicts the linear model $y = \mathbf{X}
\cdot \beta$. Thus, we seek a regularization approach that will fit a linear
model with anatomically-grouped covariates, where we expect to observe both
groupwise sparsity, where the number of groups (tract/metric combinations) with
at least one non-zero coefficients is small, as well as within-group sparsity,
where the number of non-zero coefficients within each non-zero group is small.
The sparse group lasso (SGL)
is a penalized regression technique that satisfies exactly these
criteria\cite{simon2013sparse}. It solves for a coefficient vector
$\widehat{\beta}$ that satisfies
\begin{equation}
    \widehat{\beta} = \min_\beta \frac{1}{2}
    \norm*{y - \displaystyle \sum_{\ell = 1}^{G}
    \mathbf{X}^{(\ell)} \cdot \beta^{(\ell)}}_2^2
    + \lambda_1 \displaystyle \sum_{\ell = 1}^{G}
    \sqrt{p_\ell} \norm*{\beta^{(\ell)}}_2
    + \lambda_2 \norm*{\beta}_1,
    \label{eq:sgl}
\end{equation}

@@

For given values of $\lambda_1$ and $\lambda_2$, the cost function in equation
\ref{eq:sgl} can be optimized using proximal gradient descent methods
\cite{parikh2014proximal} here implemented as a custom proximal operator that is
then optimized using the C-OPT library\cite{copt}. This supplies an estimate
of the optimal $\widehat{\beta}$ given a particular set of values for the
regulation parameters $\lambda_1$ and $\lambda_2$. We then find the optimal
values for $\lambda_1$ and $\lambda_2$ using the hyperopt library's \verb|fmin|
function\cite{Bergstra_2015} with a tree of Parzen estimators search
algorithm\cite{bergstra2011algorithms}. For the WH dataset, where the target
variable is continuous numerical, \verb|fmin| searches for metaparameters that
minimize the median absolute error, though in general a user may also
choose to minimize the root-mean-square-error or to maximize the
coefficient of determination ($R^2$). For the ALS dataset, where the
target variable is binary categorical, \verb|fmin| seeks to maximize the
classification accuracy, though in general the user may also choose to
maximize the area under the receiver operating curve (ROC AUC) or the
average precision. \adam{Enough detail provided?}

To objectively evaluate the model and guard against over-fitting,
we used a nested cross-validation scheme, depicted in
Fig~\ref{fig:cross-val} for the categorical classification case. The
input data is shuffled and then decomposed into $k$ batches, hereafter
referred to as folds. For the ALS dataset, $k=10$, while for the WH
dataset, $k=5$. For each unique fold, we hold that fold out as the
test\textsubscript{outer} set and let the remaining data comprise
the train\textsubscript{outer} set, with the subscript indicating
the depth of the nested decomposition. We further decompose each
train\textsubscript{outer} set into three folds, and again for each
unique fold, we hold out that fold as the test\textsubscript{inner}
set and let the remaining data comprise the train\textsubscript{inner}
set. At level-1 of the decomposition, we fit an SGL model using
arbitrary fixed regularization parameters $\lambda_1$ and $\lambda_2$,
training the model using train\textsubscript{inner} and evaluating
the fit using the appropriate metric on test\textsubscript{inner}.
The optimal regularization parameters are determined using the
aforementioned hyperopt library at level-0 of the decomposition. We
find optimal regularization parameters and $\widehat{\beta}$ for each
train\textsubscript{outer} set and then use those to predict values for
data in test\textsubscript{outer}. Thus each subject in the dataset has
a predicted phenotype derived from a model that never saw its particular
subject's data.

The above procedure describes $k$-fold cross validation. In fact,
we use repeated $k$-fold cross validation on the outer level of the
decomposition, so that the input data is decomposed into $k$ folds,
three times. Thus, each subject has three predicted phenotypes. We then
take the mean predicted value to summarize the skill of the model.
In the classification case, the shuffling into folds is stratified such
that each fold has a population that preserves the percentage of each
class found in the larger input data.
\adam{Clear as mud?}.

@@

>>>>

\item Diffusion MRI data from a previous study of properties of the white matter
across the lifespan\cite{yeatman2014lifespan}, containing dMRI data from 76
subjects with ages 6-50. These data were measured in a GE Discovery 750 3T MRI
scanner at the Stanford Center for Cognitive and Neurobiological Imaging. The
Stanford University IRB approved the procedures of this study. Informed consent
was obtained from each adult participant, and assent for participation was
provided by parents/guardians for children. Voxel resolution was
\num{2x2x2}$mm^3$ with 96 non-colinear directions measured with a $b=2000$
$\frac{sec}{mm^2}$ and 30 non-colinear directions measured with a $b=1000$
$\frac{sec}{mm^2}$. Data was preprocessed to correct for subject motion and the
diffusion tensor model \cite{basser1994mr} was fit in every voxel, using a
robust fit \cite{chang2005restore}. These data were acquired using a dual spin
echo sequence, in which there is sufficient time for eddy currents to subside
between the application of the gradients and the image acquisition, so no eddy
current correction was appliced. We will refer to this dataset as WH.

\item Diffusion MRI from from a previous study of the corticospinal tract
profile in patients with amyotrophic lateral sclerosis
(ALS)\cite{sarica2017corticospinal}, containing data from 24 ALS patients and 24
demographically matched healthy controls. These data were measured in a GE
Discovery 750 3T MRI scanner at the Institute of Bioimaging and Molecular
Physiology in Catanzaro. Informed consent was provided as approved by the
Ethical Committee of the University ``Magna Graecia'' of Catanzaro. Voxel
resolution was \num{2x2x2} $mm^3$ and 27 non-colinear directions were
measured with a $b=1000$ $\frac{sec}{mm^2}$. Here, both motion correction and
eddy current correction were applied before modeling the diffusioon tensor in
each voxel. We will refer to this dataset as ALS.

@@

Here, we propose that in addition to regularizing the coefficients in $\beta$,
we can also capitalize on our knowledge of the group structure of the tract
profile features in $\mathbf{X}$. The tract-metric combinations form a natural
grouping. For example, we expect that MD features within the left arcuate
fasciculus will co-vary across individuals. Likewise for FA values within the
right corticospinal tract (CST) and so on. This group structure is represented
in Fig~\ref{fig:group-structure}, which depicts the linear model $y = \mathbf{X}
\cdot \beta$. Thus, we seek a regularization approach that will fit a linear
model with anatomically-grouped covariates, where we expect to observe both
groupwise sparsity, where the number of groups (tract/metric combinations) with
at least one non-zero coefficients is small, as well as within-group sparsity,
where the number of non-zero coefficients within each non-zero group is small.
The sparse group lasso (SGL)
is a penalized regression technique that satisfies exactly these
criteria\cite{simon2013sparse}. It solves for a coefficient vector
$\hat{\beta}$ that satisfies
\begin{equation}
    \hat{\beta} = \min_\beta \frac{1}{2}
    ||y - \displaystyle \sum_{\ell = 1}^{G}
    \mathbf{X}^{(\ell)} \cdot \beta^{(\ell)}||_2^2
    + \lambda_1 \displaystyle \sum_{\ell = 1}^{G}
    \sqrt{p_\ell} ||\beta^{(\ell)}||_2
    + \lambda_2 ||\beta||_1,
    \label{eq:sgl}
\end{equation}

@@

@@

For given values of $\lambda_1$ and $\lambda_2$, the cost function in equation
\ref{eq:sgl} can be optimized using proximal gradient descent methods
\cite{parikh2014proximal} here implemented as a custom proximal operator that is
then optimized using the C-OPT library\cite{copt}. This supplies an estimate of
the optimal $\hat{\beta}$ given a particular set of values for the
meta-parameters $\lambda_1$ and $\lambda_2$.

To objectively evaluate the model and guard against over-fitting, we used a
nested cross-validation scheme, depicted in Fig~\ref{fig:cross-val} for the
categorical classification case. The input data is shuffled and then decomposed
into $k$ batches, hereafter referred to as folds. For the ALS dataset we used
$k=10$ and for the WH dataset $k=5$. For each unique fold, we hold that fold out
as the test\textsubscript{outer} set and let the remaining data comprise the
train\textsubscript{outer} set, with the subscript indicating the depth of the
nested decomposition. We further decompose each train\textsubscript{outer} set
into three folds, and again for each unique fold, we hold out that fold as the
test\textsubscript{inner} set and let the remaining data comprise the
train\textsubscript{inner} set. At level-1 of the decomposition, we fit an SGL
model using fixed regularization meta-parameters $\lambda_1$ and $\lambda_2$,
training the model using train\textsubscript{inner} and evaluating the fit on
test\textsubscript{inner}. We find the optimal values for $\lambda_1$ and
$\lambda_2$ using hyperoptimizatin. This is implemented using the hyperopt
library's \verb|fmin| function\cite{Bergstra_2015} with a tree of Parzen
estimators search algorithm\cite{bergstra2011algorithms}. For continuous
numerical $y$, \verb|fmin| searches for meta-parameter values that minimize the
median absolute error. This can also be done minimizing the root of the mean
squared error (RMSE) or to maximizing the coefficient of determination ($R^2$).
For binary categorical $y$ \verb|fmin| seeks to maximize the classification
accuracy. This can also be done maximizing the area under the receiver operating
curve (ROC AUC) or the average precision. Using hyperoptimization, we find
optimal regularization parameters and $\hat{\beta}$ for each
train\textsubscript{outer} set and then use those to predict values for data in
test\textsubscript{outer}. Thus each subject in the dataset has a predicted
phenotype derived from a model that never saw its particular subject's data.

The above procedure describes $k$-fold cross validation. In fact, we use
repeated $k$-fold cross validation on the outer level of the decomposition, so
that the input data is decomposed into $k$ folds, three times. Thus, each
subject has three predicted phenotypes. We then take the mean predicted value to
summarize the performance of the model. In the classification case, the
shuffling into folds is stratified such that each fold has a population that
preserves the percentage of each class found in the larger input data.

### 2019-12-06 reports/manuscript/sections/results.tex
Convert to docx using pandoc. Simplify some of the math used to allow that.

We also find that mislabelled subjects correspond to outlying features in the
CST. Fig~\ref{fig:class-errors} depicts the group FA values along with
FA values of mislabelled subjects, two false negatives and one false-positive.
The false negative classifications have high FA in one of the two sections of
the CST where $\norm{\widehat{\beta}} > 0$ in Fig~\ref{fig:class-profiles}.
The false positive subject has an FA that oscillates between the two group
means. Thus, the SGL method fails comprehensibly.

@@

\begin{figure}[h]
    \centering
        \includegraphics[width=0.95\textwidth]{classification_subjects_profiles.pdf}
    \caption{{\bf Model mis-classifications correspond to features identified by
       the model}. The FA in CST of individuals that are mis-classified by the
       model is compared the group FA (with shaded are indicating standard error
       of the mean). False negative classifications (individuals that have ALS,
       but are classified as patients) correspond to high FA either in one of
       the two regions of large $\norm{\widehat{\beta}}$ in
       Fig~\ref{fig:class-profiles}. The false positive classification has
       an FA profile that oscillates between the two group means.
   }
    \label{fig:class-errors}
\end{figure}

>>>>

We also find that mislabelled subjects correspond to outlying features in the
CST. Fig~\ref{fig:class-errors} depicts the group FA values along with
FA values of mislabelled subjects, two false negatives and one false-positive.
The false negative classifications have high FA in one of the two sections of
the CST where $||\hat{\beta}|| > 0$ in Fig~\ref{fig:class-profiles}.
The false positive subject has an FA that oscillates between the two group
means. Thus, the SGL method fails comprehensibly.

@@

\begin{figure}[h]
    \centering
        \includegraphics[width=0.95\textwidth]{classification_subjects_profiles.pdf}
    \caption{{\bf Model mis-classifications correspond to features identified by
       the model}. The FA in CST of individuals that are mis-classified by the
       model is compared the group FA (with shaded are indicating standard error
       of the mean). False negative classifications (individuals that have ALS,
       but are classified as patients) correspond to high FA either in one of
       the two regions of large $||\hat{\beta}||$ in
       Fig~\ref{fig:class-profiles}. The false positive classification has
       an FA profile that oscillates between the two group means.
   }
    \label{fig:class-errors}
\end{figure}
